Finalised versions of normalisation and threshold

### 2020-07-27 simulations_MMPD6.Rmd
updating documentation of MMPD

Even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption. Hence, there is a need to rank the harmonies basis how well they capture the variation in the measured variable and additionally reduce the number of harmonies for further exploration/visualization. Assuming a numeric response variable, our graphics are displays of distributions compared across combinations of categorical variables, one placed at x-axis and the other on the facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data. Thus, the idea here is to rate a harmony pair higher if this variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.

@@

One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable. We do this through Jensen-Shannon distance which is based on Kullback-Leibler divergence. Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth. We call this measure of variation as  Median Maximum Pairwise Distances (MMPD)

# Normalize distances

The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the harmonies. But maximum is not robust to the number of levels and is higher for harmonies with higher levels. Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels. The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re- normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.Suppose $d_{1},d_{2}\ldots ,d_{n}$ be a sequence of independent and identically-distributed pairwise distances and $M_{n}=\max\{d_{1},\ldots ,d_{n}\}$. Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{n}, b_{n})$ exists such that each $a_{n}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family. The normalizing constants $(a_{n}, b_{n})$ vary depending on the underlying distribution of the pairwise distances. Hence to normalize appropriately, it is important to assume a distribution of these distances.

@@

@@

@@

# Choose thresholds for harmonies

Permutation test:

Cons: Only one threshold for all harmony pairs might not be an appropriate measure but a good benchmark.

>>>>

Even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption. Hence, there is a need to rank the harmonies basis how well they capture the variation in the measured variable and additionally reduce the number of harmonies for further exploration/visualization. Assuming a numeric response variable, our graphics are displays of distributions compared across combinations of categorical variables, one placed at x-axis and the other on the facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data. Here, we have two main objectives:

- To choose harmonies for which distributions of categories are significantly different 
- To rank the selected harmonies from highest to lowest variation in the distribution of their categories. The idea here is to rate a harmony pair higher if this variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.

@@

One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable. We do this through Jensen-Shannon distance which is based on Kullback-Leibler divergence.

The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
$$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$. Probability distributions are estimated through quantiles instead of kernel density so that there is minimal dependency on selecting kernel or bandwidth.

We call this measure of variation as  Median Maximum Pairwise Distances (MMPD).

## Normalize distances

The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the harmonies. But maximum is not robust to the number of levels and is higher for harmonies with higher levels. Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels. The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re- normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.

More formally, $d_{1},d_{2}\ldots ,d_{n}$ be a sequence of independent and identically-distributed pairwise distances and $M_{n}=\max\{d_{1},\ldots ,d_{n}\}$. Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{n}, b_{n})$ exists such that each $a_{n}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family. The normalizing constants $(a_{n}, b_{n})$ vary depending on the underlying distribution of the pairwise distances. Hence to normalize appropriately, it is important to assume a distribution of these distances.

@@

# Choose thresholds for harmonies through permutation test

Cons: Only one threshold for all harmony pairs means we are assuming distribution of all harmonies pairs are similar, which might not be the case.But nevertheless, it is a good benchmark.

@@

## Minimal example

Consider three cyclic granularities $A$, $B$ and $C$ with $2$, $3$ and $4$ categories. Thus, the harmony table consisting of all possible harmony pairs (assuming all pairs are harmonies), would look like the following:

```{r harmony_min}
# harmonies <- tibble::tibble(facet_variable = c("A", "B", "A", "C", "B", "C"),
#                             x_variable  = c("B","A", "C", "A", "C", "B"),
#                             facet_levels = c(2, 3, 2, 4, 3, 4),
#                             x_levels = c(3, 2, 4, 2, 4, 3))

harmonies <- tibble::tibble(facet_variable = c("A", "B"),
                            x_variable  = c("B","A"),
                            facet_levels = c(2, 3),
                            x_levels = c(3, 2))

harmonies %>% knitr::kable()
```

### Scenario 1: Simulated same normal distributions for all combinations of categories for all rows in the harmony table.

@@
